Towards Patient-Specific Neural Field Modeling using Micro-Electrode Data

### 2011-02-06 ltx/data_paper/DataPaper.tex
added abit to estimation algorithm

As a result of the model reduction via the basis function decomposition and the decomposition of the connectivity kernel the state space model is formed
\begin{align}
 \mathbf x_{t+1}&=q(\mathbf x_t)\boldsymbol\theta+\xi \mathbf x_t+\boldsymbol e_t(\mathbf r) \\
 \mathbf y_t&=\mathbf C \mathbf x_t+\boldsymbol \epsilon_t,
\end{align}
where the nonlinear function, $q(\cdot)$, is defined as
\begin{equation}\label{eq:QmatrixForSigmapoints}
	q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'.
\end{equation}
\section{Estimation Algorithm}
Here we describe the problem of estimating non-linear state-space neural field model using the expectation-maximisation (EM) technique. The resulting estimation process is an iterative algorithm that converges on the maximum-likelihood estimates of the parameters. At each iteration we maximise an approximate lower bound on the log likelihood function. The approximate lower bound can be computed by taking the first-order truncated Taylor-series expansion of the non-linear component around the smoothed state estimate at each time instant.
\subsection{E-Step}
The unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} was used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. An additional step is required in the $\mathbf E$-step to calculate the cross-covariance matrix $\mathbf M_{t+1}=\textrm{cov}(\mathbf x_t,\mathbf x_{t+1})$. An additional element to the URTSS algorithm was also created to approximate $q(\mathbf{x}_t)$, which is required for the M-step.

>>>>

As a result of the model reduction via the basis function decomposition and the decomposition of the connectivity kernel the state space model is formed
\begin{align}
 \mathbf x_{t+1}&=q(\mathbf x_t)\boldsymbol\theta+\xi \mathbf x_t+\boldsymbol e_t(\mathbf r) \\
 \mathbf y_t&=\mathbf C \mathbf x_t+\boldsymbol \epsilon_t,
\end{align}
where the nonlinear function, $q(\cdot)$, is defined as
\begin{equation}\label{eq:QmatrixForSigmapoints}
	q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'.
\end{equation}
\section{Estimation Algorithm}
Here we describe the problem of estimating non-linear state-space neural field model using the expectation-maximisation (EM) technique. The resulting estimation process is an iterative algorithm that converges on the maximum-likelihood estimates of the parameters. In the E-step we estimate the states $\mathbf x_t$, of the systems and in the M-step we estimate the connectivity kernel parameters $\boldsymbol \theta$, disturbance variance $\sigma_d$ and the observation noise variance $\sigma_{\epsilon}$. Using the state estimates at each iteration we maximise an approximate lower bound on the log likelihood function. The approximate lower bound can be computed by taking the first-order truncated Taylor-series expansion of the non-linear component around the smoothed state estimate at each time instant.
\subsection{E-Step}
The unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} was used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. An additional step is required in the $\mathbf E$-step to calculate the cross-covariance matrix $\mathbf M_{t+1}=\textrm{cov}(\mathbf x_t,\mathbf x_{t+1})$. An additional element to the URTSS algorithm was also created to approximate $q(\mathbf{x}_t)$, which is required for the M-step.
